Self-sustained carbon monoxide oxidation oscillations on size-selected platinum nanoparticles at atmospheric pressure

### 2012-04-18 paper.tex
Updated the Lynx-hare references and fixed a few encoding errors

biological examples is the oscillation in the lynx-hare population as
reported through the number furs delivered to the Hudson Bay Company A).
This was for some time was believed to be forced by the sunspot oscillations,
however, this proved to be a coincidence B) and could be described alone
by a close dynamical coupling in between the two populations. Similar very
interesting observations have been made for our current climate changes
\cite{Svensmark1997,Svensmark1998} suggesting that the sunspot activity could
influences the level of cosmic rays and thus the cloud formation, but this
currently a topic of controversy \cite{Lockwood2007}. A beautiful example of
self-sustained oscillatory behavior in chemistry is the colorful
Belousov-Zhabotinsky reaction which is rather easy to demonstrate for an

@@

catalyst of highly dispersed Pt on γ−Al2O3 \cite{BEUSCH1972}. To gain a deeper
insight into the fundamental aspects of such oscillatory reactions this work
was followed up by Ertl on Pt(110) surfaces under well-defined conditions
\cite{EISWIRTH1986}. Ertl demonstrated how the switching in between the
unreconstructed Pt(110) surface and the reconstruction 2x1-Pt(110) surface
combined with big changes of oxygen sticking coefficient on the two surfaces
could drive the self-sustained oscillations. This and many other interesting
aspects of self-sustained and forced oscillatory behavior and spatio–temporal
self-organization were investigated by Ertl and coworkers on well-defined
single crystals see for example reviews \cite{IMBIHL1995,JAKUBITH1990} and in
particular the Nobel Lecture \cite{Ertl2008}. One of the reasons this has
attracted so much interest is naturally the importance of such catalytic
reactions and therefore a demand to gain a better understanding of their
behavior with the goal of further improvement. CO oxidation is one of the most
studied reactions due to its importance in both automotive catalysis and in
the removal of trace CO in hydrogen streams \cite{IbsBog}. The oscillatory
behavior is, however, limited to either Pt or the CO oxidation, but
have been found for many other metallic catalyst like on for example Rh, Ir,
and Pd and for oxidations reactions in general as well as for NO and CO
reactions and even for hydrogenation reactions \cite{Lund2000,SALES1982}. Most
of the studies are made on single crystal or poly crystalline films or
filaments, mainly in conjunction with ultra-high vacuum equipment so that the
surfaces can be characterized using the surface science methods. The amount of
work described on supported catalyst seems more limited obviously because it
is so much more difficult to characterize those systems and because additional
complexities are present such as mass and heat transport. There even exist
reports in the literature of how large scale facilities such as an industrial
ammonia fixed-bed synthesis reactor in Germany in 1989 started oscillating for
several hours. This could be modeled by a coupling of the exothermic process
and the heat management in the reactor system \cite{Machado2005}, but one
cannot rule out that the observed oscillations could in some systems also be
created by inherent processes on the catalyst itself. In general such
oscillations would be considered highly undesirable in industry since they
could be difficult to control and result in run-away incidents in particular
for exothermic reactions resulting in damaged catalyst due to sintering at
higher temperatures. On the other hand there exist examples in the literature
where forced oscillations seem to improve the overall activity
\cite{IMBIHL1995,Machado2005} . One particular interesting process here could
be the methanol synthesis where reducing the CO$_2$ contents for a shorter period
of time result in a subsequent higher methanol rate over compensation for the
lack of methanol production when running in the reducing H$_2$ and CO alone
\cite{Dynamics-Bog}. This enhancement can be explained by extensive wetting
and un-wetting of the active Copper catalyst dependent on the oxidation
potential \cite{Hansen2002} i.e. the active area is strongly increased when
reduced but slowly returns to the less active state when the CO$_2$ is present.

@@

A) D. A. McLulich, Variations in the Number of Varying Hare,University of Toronto
Press, Toronto, 1937. (we need to dig in for the correct ref here),
B) I remember reading of this in a biology book back in the seventies, check)

>>>>

biological examples is the oscillation in the lynx-hare population as reported
through the number furs delivered to the Hudson Bay Company
\cite{MacLulich,Elton1942}). This was for some time was believed to be forced
by the sunspot oscillations \cite{Elton1924}, however, even though the
realtionship has been debated even as recently as 1993\cite{SINCLAIR1993} the
concensuss seems to be that this is not the case \cite{Elton1942,Kielland2010}
and the periodic cycle can be explained by more classic wheather phenomena..
Similar very interesting observations have been made for our current climate
changes \cite{Svensmark1997,Svensmark1998} suggesting that the sunspot activity
could influences the level of cosmic rays and thus the cloud formation, but
this currently a topic of controversy \cite{Lockwood2007}. A beautiful example
of self-sustained oscillatory behavior in chemistry is the colorful Belousov-
Zhabotinsky reaction which is rather easy to demonstrate for an

@@

catalyst of highly dispersed Pt on $\gamma$−Al$_2$O$_3$ \cite{BEUSCH1972}. To
gain a deeper insight into the fundamental aspects of such oscillatory
reactions this work was followed up by Ertl on Pt(110) surfaces under well-
defined conditions \cite{EISWIRTH1986}. Ertl demonstrated how the switching in
between the unreconstructed Pt(110) surface and the reconstruction 2x1-Pt(110)
surface combined with big changes of oxygen sticking coefficient on the two
surfaces could drive the self-sustained oscillations. This and many other
interesting aspects of self-sustained and forced oscillatory behavior and
spatio--temporal self-organization were investigated by Ertl and coworkers on
well-defined single crystals see for example reviews
\cite{IMBIHL1995,JAKUBITH1990} and in particular the Nobel Lecture
\cite{Ertl2008}. One of the reasons this has attracted so much interest is
naturally the importance of such catalytic reactions and therefore a demand to
gain a better understanding of their behavior with the goal of further
improvement. CO oxidation is one of the most studied reactions due to its
importance in both automotive catalysis and in the removal of trace CO in
hydrogen streams \cite{IbsBog}. The oscillatory behavior is, however, limited
to either Pt or the CO oxidation, but have been found for many other metallic
catalyst like on for example Rh, Ir, and Pd and for oxidations reactions in
general as well as for NO and CO reactions and even for hydrogenation reactions
\cite{Lund2000,SALES1982}. Most of the studies are made on single crystal or
poly crystalline films or filaments, mainly in conjunction with ultra-high
vacuum equipment so that the surfaces can be characterized using the surface
science methods. The amount of work described on supported catalyst seems more
limited obviously because it is so much more difficult to characterize those
systems and because additional complexities are present such as mass and heat
transport. There even exist reports in the literature of how large scale
facilities such as an industrial ammonia fixed-bed synthesis reactor in Germany
in 1989 started oscillating for several hours. This could be modeled by a
coupling of the exothermic process and the heat management in the reactor
system \cite{Morud1998}, but one cannot rule out that the observed oscillations
could in some systems also be created by inherent processes on the catalyst
itself. In general such oscillations would be considered highly undesirable in
industry since they could be difficult to control and result in run-away
incidents in particular for exothermic reactions resulting in damaged catalyst
due to sintering at higher temperatures. On the other hand there exist examples
in the literature where forced oscillations seem to improve the overall
activity \cite{IMBIHL1995,Machado2005} . One particular interesting process
here could be the methanol synthesis where reducing the CO$_2$ contents for a
shorter period of time result in a subsequent higher methanol rate over
compensation for the lack of methanol production when running in the reducing
H$_2$ and CO alone \cite{Dynamics-Bog}. This enhancement can be explained by
extensive wetting and un-wetting of the active Copper catalyst dependent on the
oxidation potential \cite{Hansen2002} i.e. the active area is strongly
increased when reduced but slowly returns to the less active state when the
CO$_2$ is present.

@@

B) I remember reading of this in a biology book back in the seventies, check)
